---
title: "HW2-EDA with R - Part1: Plotting"
author: "misken"
date: "January 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1 - Familiarize yourself with the data and the assignment

In this assignment you'll do some exploratory data analysis
with R on a dataset US census information based on the [American Community Survey](https://www.census.gov/programs-surveys/acs/). The version of
this data that we'll be using is from a Kaggle dataset. See
https://www.kaggle.com/muonneutrino/us-census-demographic-data/data and look
at the column metadata for the county level table to see what the various
columns mean.

As we did in class, you'll be creating an R Markdown document to
both prepare a dataset and do some analysis as well as to document the
steps you did (and answer some questions I'll throw at you).

You'll notice a few "Hacker Extra" tasks
thrown in. These are for those of you who want to go a little above and beyond
and attempt some more challenging tasks. And, feel free to do a some free form
Hacker Extra style work yourself - in other words, do something beyond what
was asked. You'll learn more.

## Step 2 - Create a new R Markdown document

Save this file as a new R Markdown document and name it **HW2_EDAwR_[_your last name_].Rmd**. Mine would
be called **HW2_EDAwR_isken.Rmd**. Save it into the same folder as this file. 

This will be your working directory. 

Create a new R Project based on this working directory.

**Hacker Extra** - Put this project folder under version control using git. Only
track the .R file, the .Rmd file and your .gitignore file. Create a remote repo
for this homework assignment in your GitHub account. Periodically push your commits
to GitHub. When done with the assignment, paste in the URL of your GitHub project
so that I can go look at it.

## Step 3 - Set working directory and explore data folder

Set the working directory or create an R Studio project in the current folder. You'll notice that there is a folder named **data**.
Inside of it you'll find the data files for this assignment:

- **acs2015_census_tract_data.csv**
- **acs2015_county_data.csv**
- **us_census_bureau_regions_and_divisions.csv**
- **county.csv**
- **county_MI.csv**

The first two files are from the Kaggle site for this dataset. The two files,
**county.csv** and **county_MI.csv**, are files that I created. The first
contains county level data from all states and the other just data from
Michigan. The R script, **census_data_prep.R** was used to create these files
and to merge information from the **us_census_bureau_regions_and_divisions.csv**
file with the county data.

You might want to use shell commands or a text editor like Geany to snoop the files a bit.


## Step 4 - Complete the following R tasks and answer questions

Now you'll need to complete the following tasks in R. Just like we did in class, you should
use a combination of markdown text (be concise, no need to write tons of text) 
to explain what you are doing and R code chunks to actually do it.
When you are done, use the "Knit" button to generate an HTML file from your R Markdown.
You'll be submitting BOTH the completed R Markdown file as well as the generated HTML file. You'll notice that I've left some empty code chunks
as well as some partially filled in code chunks. Of course, feel free to
add more code chunks as needed.

Let's load a few libraries we'll need:

```{r}
library(dplyr)
library(ggplot2)
```


### Problem 1 - Get the data into a dataframe

You should read the **county.csv** file into a data frame named county_df and
**county_MI.csv** into a data frame named county_MI.

```{r read_countydata}
county_df <- read.table("data/county.csv", header=TRUE, sep=",", stringsAsFactors=FALSE)
county_MI <- read.table("data/county_MI.csv", header=TRUE, sep=",", stringsAsFactors=FALSE)
```

Check the structure of each data.frame and use the `summary` function to begin
exploring the data. Are there any NA values? What 
differences do you notice between the structure of the two data frames.
# Structure of Data Frame county_df
```{r}
str(county_df)
```

# Structure of Data Frame county_MI
```{r}
str(county_MI)
```

The county_MI data frame contains data for Michigan only and includes an
peninsula field. See **census_data_prep.R** to see how this field was
created. It involves scraping a table out of a web page.

Unless otherwise specified, you should use the ggplot2 package for all of the
following plots.

### Problem 2 - Basic plots for Income

Create a histogram, boxplot, kernel density plot and violin plot for median
income for the entire US. 

Write a short paragraph describing the distribution of median incomes.

From the Histogram as we can see that both the Median and the Mean incomes are equal and hence the distribution of income over the entire US is symmetrically distributed. 

From the Boxplot, it can be said that, the Median Income for the entire US is closer to the lower end of the spectrum of all the values and is closer to the end of the left whiskers (down whiskers) than to that of the end of the right whisker (top whiskers). The Median income for the entire US is along the mid line of the box, starting a little below 50,000 and we can approximate the same as around $ 45,000. 

From both the Kernel Density Plot as well as the Violin plot we can see that the spread of the Income for the entire US is topping and buldging respectively somehwere around 45,000.

# Histogram

```{r}
median_income <- median(county_df$Income, na.rm = TRUE)
median_income

mean_income <- median(county_df$Income, na.rm = TRUE)
mean_income

library(ggplot2)
ggplot(data = county_df) + 
  geom_histogram(aes(x = Income), binwidth = 3000, fill = "steelblue", colour="Black")

```

# BoxPlot 
There needs to be some value corresposnding to x aesthetic, so used 1. 

```{r}
ggplot(county_df, aes(y=Income, x=1)) + geom_boxplot(fill = "orange", alpha = 0.2)

```

# Kernel Density Plot

```{r}
library(ggplot2)
ggplot(county_df, aes(x=Income)) + geom_density(fill = "orchid", alpha = 0.2)
```

# Violin Plot

```{r}
ggplot(county_df, aes(x="US", y=Income)) + geom_violin(fill = "red", alpha = 0.2)
```

### Problem 3 - Histogram of total population

Create a histogram of total population for counties in the US.

```{r histo_totpop}
ggplot(data = county_df) + geom_histogram(aes(x = TotalPop))
```

Repeat the histogram about but make the following style changes to the plot.

* Add a title such as "Distribution of population for US counties"
* Change y-axis label to "Number of counties"
* Change x-axis label to "Total Population"
* Change the colors of the bars to something other than the default black

```{r histo_stylechanges}
library(ggplot2)
ggplot(data = county_df) + geom_histogram(aes(x = TotalPop), fill = "violet", color = "black") + labs(
    x = "Total Population",                                    # x axis title
    y = "Number of counties",                                  # y axis title
    title = "Distribution of population for US counties"       # main title of figure
  ) + scale_x_log10()
```

**Hacker Extra** - Also create a kernel density plot overlaid on a histogram of
Income. 

```{r}
ggplot(county_df, aes(x=Income, y=..density..)) + geom_histogram(fill="cornsilk", color="grey60", size=.2) + geom_density()
```


Repeat the histogram but only for counties with less than 2.5 million people.

```{r histo_lt2.5m}
# Created a dataframe with less than 2.5 million people
county_sub <-subset(county_df, TotalPop < 2500000)
county_sub

# Calling data from the new data frame, county_sub with less than 2.5 million people
library(ggplot2)
ggplot(data = county_sub) + geom_histogram(aes(x = TotalPop), fill = "orange", color = "black") + labs(
    x = "Total Population",                                                                      # x axis title
    y = "Number of counties",                                                                    # y axis title
    title = "Distribution of population for US counties with less than 2.5 million poeple"       # main title of figure
  )

```


### Problem 4 - How do Michigan's peninsulas compare?

Create a grouped box plot showing Income by peninsula. Make sure
your graph has a title and good axis labels.

```{r bp_inc_by_peninsula}
ggplot(county_MI, aes(x=peninsula, y=Income, color=peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                    # x axis title
    y = "Income",                                       # y axis title
    title = "Distribution of Income by Peninsula"       # main title of figure
  )
```

Do the peninsulas differ with respect to distribution of median income over the counties?

Yes, from the Boxplot as shown above, it looks like there is a significant difference in the distribution of median income in the peninsulas over the counties.


Create several more boxplots showing how the peninsulas compare with respect to
some of the other variables. Summarize your findings in a few short paragraphs.

# i) Population Distribution Across Peninsula

The total population distribution for LP is higher than that of the UP. The Median of the Total Population distribution for Upper Peninsula is closer to the lower end of the spectrum for all of the population value and is closer to the end of the left whiskers (down whiskers) than to that of the end of the right whisker (top whiskers), indicating thin population distribution in Upper Peninsula as compared to that of the Lower Peninsula. On the other hand, the Median of the Total Population distribution is higher in Lower Peninsula as compared to that of the Upper Peninsula.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=TotalPop, fill=peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                    # x axis title
    y = "Total Population",                             # y axis title
    title = "Population Distribution by Peninsula"      # main title of figure 
) + scale_y_log10()
```

# ii) Poverty Distribution across Peninsula

The Median of the poverty distribution for the UP is somewhere around 16 and is closer to the end of the left whiskers (down whiskers) than to that of the end of the right whisker (top whiskers), indicating that the Median of the Poverty distribution is closer to the lower end of the entire spectrum in Upper Peninsula as compared to that of the Lower Peninsula. In case of the LP or Lower Peninsula, the Median of the poverty distribution is around 18 and is towards the right whiskers (top whiskers) as compared to the left whiskers (down whiskers). The Median of the poverty distribution for Lower Peninsula is at a higher side as compared to that of the Upper Peninsula.

```{r more_boxplplots}
ggplot(county_MI, aes(x=peninsula, y=Poverty, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                    # x axis title
    y = "Poverty",                                      # y axis title
    title = "Poverty Distribution by Peninsula"         # main title of figure 
) 

```

# iii) Child Poverty Distribution across Peninsula

The Median of the Child Poverty Distribution for Upper Peninsula is at 25 and it goes around the middle of the entire spectrum of the boxplot. From the Box-plot it looks like all the Quartiles for the Upper Peninsula are uniformly and equally distributed. The Median of the Child Poverty Distribution for Lower Peninsula is slightly below 25 and the Median of the Child Poverty Distribution for the entire spectrum of the boxplot is towards the top whiskers (or the right whiskers).

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=ChildPoverty, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                          # x axis title
    y = "Child Poverty",                                      # y axis title
    title = "Child Poverty Distribution by Peninsula"         # main title of figure 
) 

```

# iv) Income per capita Distribution by Peninsula

The Median of the Income Per Capita distribution across Upper Peninsula is lower than that of the Median of the Income per capita distribution in the Lower Peninsula. The Median of the Income per capita distribution for the Upper Peninsula is lower than the value of 22,500, while the Median of the Income per capita distribution for the Lower Peninsula is greater than the value of 22,500.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=IncomePerCap, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                              # x axis title
    y = "Income Per Capita",                                           # y axis title
    title = "Income per capita Distribution by Peninsula"         # main title of figure 
) 
```

# v) Employment Distribution Across Peninsula

Considering the Employment distribution across both the Upper & the Lower Peninsula, it looks like the Median of the Employment Distribution for the Upper Peninsula is much lower as compared to that of the Lower Peninsula.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=Employed, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                           # x axis title
    y = "Employment",                                          # y axis title
    title = "Employment Distribution Across Peninsula"         # main title of figure 
) + scale_y_log10()
```

# vi) Unemployment Distribution Across Peninsula

The Median of the Unemployment Distribution across Upper Peninsula looks more towards the lower end of the down whisker (left whisker), within the entire spectrum of the Unemployment Distribution in Upper Peninsula. The Median of the Unemployment Distribution for Upper Peninsula is low as compared to the Median of the Unemployment Distribution for the Lower Peninsula. However, when compared to the Unemployment Distribution with the Lower Peninsula, the distribution of Unemployment in the Upper Peninsula appears to be more on a cummulative scale.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=Unemployment, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                           # x axis title
    y = "Unemployment",                                          # y axis title
    title = "Unemployment Distribution Across Peninsula"         # main title of figure 
) 
```

# vii) Citizens Distribution Across Peninsula

The Median of the number of Citizen Distribution looks to be on the higher side for the Lower Peninsula as compared to that of the Upper Peninsula. The Median value of the Citizen Distribution is closer towards the lower end of the entire spectrum, for the Upper Peninsula. While the Median value of the Citizen Distribution across the Lower Peninsula seems to be little dragged towards the down whisker (left). The entire quartile for the citizens distribution across both the Upper Peninsula is smaller as compared to that of the Lower Peninsula.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=Citizen, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                         # x axis title
    y = "Citizen",                                           # y axis title
    title = "Citizens Distribution Across Peninsula"         # main title of figure 
) + scale_y_log10()
```

# viii) Number of Men & Women Across the Peninsula

The Median of the number of Men across the Upper Peninsula seems to be more closer towards the down whisker (left whisker). The Median of the number of Men across the Lower Peninsula seems to be little inclined towards the down whisker (left whisker).The entire quartile for the distribution of the number of Men is more in Lower Peninsula as compared to that of the Upper Peninsula.

The Median of the number of Women across the Upper Peninsula seems to be more closer towards the down whisker (left whisker). The Median of the number of Women across the Lower Peninsula seems to be little inclined towards the down whisker (left whisker).The entire quartile for the distribution of the number of Women is more in Lower Peninsula as compared to that of the Upper Peninsula.

```{r more_boxplots}
# Number of Men
ggplot(county_MI, aes(x=peninsula, y=Men, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                 # x axis title
    y = "Men",                                       # y axis title
    title = "Number of Men Across Peninsula"         # main title of figure 
) + scale_y_log10()

# Number of Women 
ggplot(county_MI, aes(x=peninsula, y=Women, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                   # x axis title
    y = "Women",                                         # y axis title
    title = "Number of Women Across Peninsula"         # main title of figure 
) + scale_y_log10()

```

# ix) Mean Commute Time Across Peninsula

The Median of the Mean Commute Time in Upper Peninsula seems to be more closer towards the Median value of the Mean Commute Time, which is approximately 17. It looks like the Mean Commute Time in Upper Peninsula is less than that of the Lower Peninsula. The Median of the Mean Commute Time in Lower Peninsula seems to be little inclined towards the down whisker (left whisker) of the entire quartile.

```{r more_boxplots}
ggplot(county_MI, aes(x=peninsula, y=MeanCommute, fill = peninsula)) + geom_boxplot() + labs(
    x = "Peninsula",                                     # x axis title
    y = "Mean Commute",                                  # y axis title
    title = "Mean Commute Time Across Peninsula"         # main title of figure 
) 
```

### Problem 5 - How do regions of the country compare?

In county_df you'll see that there's a Region field. Create a set of faceted 
histograms showing how some variable of interest to you compares by region. For 
example, you could explore how the regions compare with respect to the
distribution of MeanCommute time. Make sure your graph has a title and good axis
labels.

# Mean Commute Time Across Various Regions

```{r faceted_histo}
ggplot(county_df, aes(x=MeanCommute, fill = Region)) + geom_histogram() + labs(
    x = "Mean Commute Time",                                   # x axis title
    y = "Count",                                               # y axis title
    title = "Mean Commute Time Across Various Regions"         # main title of figure 
) + facet_wrap(~Region) 
```

# Total Population Distribution & Distribution of Population By Gender

```{r faceted_histo}
# Featured Engineering of PercentageMen & PercentageWomen
county_df$PercentageMen <- (county_df$Men/county_df$TotalPop)*100
county_df$PercentageWomen <- (county_df$Women/county_df$TotalPop)*100

# Dropping a Column added by mistake
county_df$PercentageMale <- NULL

# Total Population Distribution
ggplot(county_df, aes(x=TotalPop, fill = Region)) + geom_histogram() + labs(
    x = "Total Population",                                                # x axis title
    y = "Count",                                                           # y axis title
    title = "Total Population Distribution Across Various Regions"         # main title of figure 
) + facet_wrap(~Region) + scale_x_log10()

# Percentage of Men Across Total Population Distribution
ggplot(county_df, aes(x=PercentageMen, fill = Region)) + geom_histogram() + labs(
    x = "Men Population %",                                               # x axis title
    y = "Count",                                                          # y axis title
    title = "Men Population Distribution Across Various Regions"          # main title of figure 
) + facet_wrap(~Region) + scale_y_log10()

# Percentage of Women Across Total Population Distribution
ggplot(county_df, aes(x=PercentageWomen, fill = Region)) + geom_histogram() + labs(
    x = "Women Population %",                                              # x axis title
    y = "Count",                                                           # y axis title
    title = "Women Population Distribution Across Various Regions"         # main title of figure 
) + facet_wrap(~Region) + scale_y_log10()


```


### Problem 6 - Basic bar chart

Create a bar chart showing the number of counties by Region in the US. Make sure
your graph has a title and good axis labels.

```{r bar_counties_by_region}
library(reshape2)
ggplot(data=county_df, aes(x=Region, fill = Region)) +
    geom_bar(stat="count") + labs(
    x = "Region",                                             # x axis title
    y = "Number of Counties",                                 # y axis title
    title = "Number of Counties by Regions in the US"         # main title of figure 
) 

```

### Problem 7 - Basic scatter plots

Create a scatter plot using the Michigan only data frame comparing
Income and MeanCommute.

```{r xy_inc_commute}
ggplot(data = county_MI, aes(x = MeanCommute, y = Income)) + geom_point()
```

Create a modified version of the above scatter plot in which the point color
is mapped to peninsula.

```{r xy_inc_commute_peninsula}
ggplot(county_MI, aes(MeanCommute, Income, color = peninsula)) + geom_point()
```

Your thoughts on the relationship between these variables?

From the Scatterplots of the two variables Income & MeanCommute, we can see random scatter where each point in the data sets represent the corresponding value of the Income for a given Mean Commute Time and vice versa. It looks like the data points produce a pattern from left to right with some data points for the Lower Peninsula as Outliers. Overall, from the Sctterplots visuals it seems like the variables Income and the Mean Commute Time has a weak correlation betwen each other.

Now, create a few more scatter plots using variables of your choosing in which you facet by peninsula as well. How do the relationships compare across the peninsulas?

# i) Scatterplots of Total Population by Income Across the Peninsula

From the Scatterplots of the two variables, Total Population and the Income it looks like that there is a weak correlation between the Total Population and the Income across both the Upper and Lower Peninsula.

```{r xy_faceted_peninsula}
ggplot(county_MI, aes(Income, TotalPop, color = peninsula)) +
geom_point() + facet_grid(peninsula ~ .)
```

# ii) Scatterplots of Total Population by Citizen Across the Peninsula 

From the Scatterplots of the two variables Total Population and the Citizen, there seems to be a steady slope between these two variables Total Population and Citizen. Hence, we can say that these two variables Total Population and Citizen have a Positive Correlation or in other words these two variables have a Strong Correlation between each other across both the Upper & Lower Peninsula in Michigan.

```{r xy_faceted_peninsula}

ggplot(county_MI, aes(Citizen, TotalPop, color = peninsula)) +
geom_point() + facet_grid(. ~ peninsula)

```

# iii) Scatterplots of Total Population by Employment Across the Peninsula

From the Scatterplots of the two variables, Total Population and the Unemployment, it looks like that there is a weak correlation between the Total Population and the Unemployment across both the Upper and Lower Peninsula.

```{r xy_faceted_peninsula}
ggplot(county_MI, aes(Unemployment, TotalPop, color = peninsula)) +
geom_point() + facet_grid(peninsula ~ .)
```


Now use the full US data frame.

Create a scatter plot of TotalPop by Income.

```{r xy_pop_inc}
# Plot of Total Population by Income, Not Scaled
ggplot(data = county_df, aes(x = Income, y = TotalPop)) + geom_point() 
```

The widely varying values make this difficult to interpret in terms of looking
for patterns. So, figure out a way to either change the axis scales or 
the plotted data itself to use a base 10 logarithmic scale (or transform).

```{r}
# Scaled Plot of Total Population by Income
ggplot(data = county_df, aes(x = Income, y = TotalPop)) + geom_point() + scale_y_log10() 
```


Write a short paragraph on why one might want to use a log scale or log transform of data when plotting.

The datasets might have values that vary from each other on a large scale as seen in the first scatterplot (# Plot of Total Population by Income, Not Scaled) above between Total Population and the Income. It is very hard to get a clarity on the data points whereby the datasets are overlapping on one another making it hard to comprehend any relationship between these two variables. This property of variability between two variables of data is due to the uneven variation of values with the mean value in the datasets often known as heteroscedastic variance. Using a log scale or, equivalently, using the log transform 'fixes' the uneven variability and gives a scale that is unbounded on both ends.

Also, for time-series data, the ability of a log scale to make fractional changes is immensely powerful. With log scale a constant percentage change is seen as a constant vertical distance so a constant growth rate is seen as a straight line. That is often a substantial advantage.

Another reason when log scale proves to be powerful, when it makes more realistic to represent the data in a reciprocal manner, as for example in research areas where someone is dealing with minutely small numbers.

### Hacker Extra - More analysis

Do some additional plotting to look for interesting patterns or relationships
either in the full US dataset or the MI only dataset. In addition, try out
some of ggplots additional features such as themes, axis formatting, gridline
control, or others.

# i) Percentage of Women & Men Population Distribution For US Across Regions

The histogram for the percentage of Men and Women Population across the entire US for the Regions Midwest, Northeast, South, and East seemed to be skewed to the left and right respectively. 
```{r}
library(ggthemes)

# Percentage Distribution of Women Population Across Regions
g2 <- ggplot(county_df, aes(x=Women/TotalPop*100)) + geom_histogram() + labs(
    x = "Women Population %",                                                  # x axis title
    y = "Count",                                                               # y axis title
    title="Percenatge of Women Population Distribution Across Regions"         # main title of figure 
) + facet_wrap(~Region) 
g2 + theme_economist() + scale_colour_economist()

# Percentage Distribution of Men Population Across Regions
g2 <- ggplot(county_df, aes(x=Men/TotalPop*100)) + geom_histogram() + labs(
    x = "Men Population %",                                                   # x axis title
    y = "Count",                                                              # y axis title
    title="Percenatge of Men Population Distribution Across Regions"          # main title of figure 
) + facet_wrap(~Region) 
g2 + theme_excel() + scale_colour_excel()

```

# ii) Distribution of Total Population & Income Per Capita In Michigan

There appears to be a linearity for the two variables Total Population and Income Per Capita, in both the Lower & Upper Peninsula with some of the outliers scattered within the data sets. The outliers for the Lower Peninsula in describing the relationship between Total Population and Income Per Capita seems to be more than the outliers for the data points of Upper Peninsula.

```{r}
ggplot(county_MI, aes(IncomePerCap, TotalPop, color = peninsula)) + geom_point() + scale_y_log10() + geom_smooth(method = "lm") + labs(
    x = "Income Per Capita",                                                              # x axis title
    y = "Total Population",                                                               # y axis title
    title="Distribution of Total Population & Income Per Capita Across Peninsula"         # main title of figure 
)
```

# iii) Employment Distribution Across Total Population Across UP & LP in Michigan

There seems to have a positive correlation between Employment Distribution across the Total Population in both the Upper & Lower Peninsula in Michigan.

```{r}
ggplot(county_MI, aes(Employed, TotalPop, color = peninsula)) + geom_point() + scale_y_log10() + geom_point(aes(shape = peninsula), size = 3) + labs(
    x = "Employed",                                                              # x axis title
    y = "Total Population",                                                      # y axis title
    title="Employment Distribution in Total Population Across Peninsula"         # main title of figure 
)
```

# iv) Unemployment Distribution Across Total Population Across UP & LP in Michigan

The Unemployment Distribution with respect to the Total Population seems to have a scattered relationship amongst the datapoints within the datasets and hence from the visuals of the graphical results we can conclude that the Unemployment Distribution and the Total Population across both the Upper & Lower Peninsula has a weak correlation. 

```{r}
ggplot(county_MI, aes(Unemployment, TotalPop, color = peninsula)) + geom_point() + scale_y_log10() + geom_point(aes(shape = peninsula), size = 3) + labs(
    x = "Unemployment",                                                            # x axis title
    y = "Total Population",                                                        # y axis title
    title="Unemployment Distribution in Total Population Across Peninsula"         # main title of figure 
)
```

# v) Men Population Distribution in US Across Regions

The Median of the percentage of Men Population Distribution across the various Regions such as the Midwest, Northeast, South, and West is in and around 50, with the upper whisker's outliers extending almost near 70 for South & West Region. And the lower whisker's outliers range extending almost near to 40 for South Region.

```{r}
ggplot(county_df, aes(x=Region, y=PercentageMen, fill = Region)) + geom_boxplot() + labs(
    x = "Region",                                                    # x axis title
    y = "Men Percentage",                                            # y axis title
    title = "Men Population Distribution by Regions in US"           # main title of figure 
) 
```

# vi) Women Population Distribution in US Across Regions

The Median of the percentage of Women Population Distribution across the various Regions such as the Midwest, Northeast, South, and West is in and around 50, with the upper whisker's outliers extending almost near to 60 for South & near to 55 for the West Region. And the lower whisker's outliers range extending almost near to 30 for both the South & West Region.

```{r}
ggplot(county_df, aes(x=Region, y=PercentageWomen, fill = Region)) + geom_boxplot() + labs(
    x = "Region",                                                    # x axis title
    y = "Women Percentage",                                          # y axis title
    title = "Women Population Distribution by Regions in US"         # main title of figure 
)
```

